---
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::word_document2: default
  bookdown::html_document2: default
documentclass: book
bibliography: references.bib
---

```{r echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
library(randomForest)
library(ggplot2)

tidy_ikea # make sure that tidy.R is executed and tidy_ikea is loaded into memory

###############
## ANALYSIS ###
###############

n_trees <- 10

#TODO: set to 2000 before handing in

# correlation price + old_price

cor.test(tidy_ikea$price_eur, tidy_ikea$old_price_eur, 
         method = "pearson")

# importance - drop_na

rf_ikea <- tidy_ikea %>%
  select(-old_price_eur) %>%
  drop_na()

rf_ikea$designer <- fct_lump_n(rf_ikea$designer, 50)
rf_ikea$name <- fct_lump_n(rf_ikea$name, 49)

result_rf <- randomForest(price_eur ~ ., rf_ikea, ntree=n_trees,
                          keep.forest=FALSE, importance=TRUE)
imp_drop_na <- importance(result_rf)

# importance - dummy value instead of na

rf_ikea <- tidy_ikea 

rf_ikea$size_m3[is.na(rf_ikea$size_m3)] <- -1000

rf_ikea <- rf_ikea %>%
  select(-old_price_eur) %>%
  drop_na()

rf_ikea$designer <- fct_lump_n(rf_ikea$designer, 50)
rf_ikea$name <- fct_lump_n(rf_ikea$name, 49)

result_rf <- randomForest(price_eur ~ ., rf_ikea, ntree=n_trees,
                        keep.forest=FALSE, importance=TRUE)
imp_size_minus_1000 <- importance(result_rf)


# importance - na roughfix

rf_ikea <- tidy_ikea %>%
  select(-old_price_eur)

rf_ikea$sellable_online <- factor(rf_ikea$sellable_online)
rf_ikea$other_colors <- factor(rf_ikea$other_colors)

rf_ikea$designer <- fct_lump_n(rf_ikea$designer, 50)
rf_ikea$name <- fct_lump_n(rf_ikea$name, 49)

result_rf <- randomForest(price_eur ~ ., rf_ikea, ntree=n_trees, importance=TRUE, na.action = na.roughfix)
imp_na_roughfix <- importance(result_rf)


# mean importance

mean_imp <- as.data.frame((imp_na_roughfix+imp_size_minus_1000+imp_drop_na)/3)

mean_imp$varnames <- rownames(mean_imp)
rownames(mean_imp) <- NULL
mean_imp <- mean_imp %>%
  rename('PercentIncMSE' = '%IncMSE')
```

# Results {#results}
\hfill\textcolor{gray}{by J. Pein}

## TODO: set n_trees to 2000 before handing in

In this chapter, the result of the analysis of the feature importance of different features on the response variable price of Ikea products are presented.

As described in section \@ref(rf), the feature importance was calculated using permuting feature importance of the _randomForest_ R package. In this analysis, feature importance is derived from the percentage increase of the mean squared error (MSE) of the overall random forest regression model in regard to the response variable `price_eur`. A larger percentage increase of the MSE implies greater feature importance. Conversely, a lower percentage increase of the MSE translates to less significant feature importance.

```{r mean-feature-importance, echo=FALSE, out.width="100%", fig.cap="A plot showing the mean feature importance of the predictor variables on the response variable"}
ggplot(mean_imp, aes(y=reorder(varnames, PercentIncMSE), x=PercentIncMSE)) + 
  geom_point() +
  scale_color_manual(values = mycolors) +
  theme_minimal() +
  labs(x = "% Increase MSE", y = "Feature", title = "Mean Feature Importance")
```


Thus, as can be seen in figure \@ref(fig:mean-feature-importance), the most important feature is `size_m3`  with an increase of the MSE of 182%. The second, third and fourth most important features are `designer` with an increase of 120%, `name` with an increase of 114% and `category` with an increase of 105%. The fifth most important feature is `other_colors` with a MSE increase of 78% and the least important feature is `sellable_online` with a 9% increase.

These results are further discussed in the following chapter.



